### 2021-06-09 notes/np_notes.tex
stash 0608
run sims for longer

It seems like we are seeing some resonances when $g_2 / g_1$ hits half-integer
multiples. For convenience, we go to the co-rotating frame with $g_1$ where we
then have
\begin{equation}
    \p{\rd{\uv{s}}{t}}_{\rm rot, 1}
        = \alpha\p{\uv{s} \cdot \uv{l}_1}\p{\uv{s} \times \uv{l}_1}
            + g_1\p{\uv{s} \cdot \uv{J}}.
\end{equation}
Here, now, the mode frequencies are $g_1 \Rightarrow 0$ and $g_2 \Rightarrow g_2
- g_1$. Decompose then $\uv{l}$ (we drop the subscript) into a mean and fluctuating
piece $\bar{\bm{l}}$ and $\bm{l}'$ (where the former is still a vector of unit
length), then:
\begin{equation}
    \p{\rd{\uv{s}}{t}}_{\rm rot, 1}
        = \alpha\p{\uv{s} \cdot \bar{\bm{l}}}\p{\uv{s} \times \bar{\bm{l}}}
            + g_1\p{\uv{s} \cdot \uv{J}}
            + \alpha\s{
                \p{\uv{s} \cdot \bar{\bm{l}}}
                \p{\uv{s} \times \bm{l}'} +
                \p{\uv{s} \cdot \bm{l}'}
                \p{\uv{s} \times \bar{\bm{l}}}
            }.
\end{equation}
This is already looking painful. Let's write down the Hamiltonian to have a
faster path to EOM in coordinates:
\begin{align}
    H_{\rm rot, 1} ={}& -\frac{\alpha}{2}\p{\uv{s} \cdot \uv{l}}^2
            - g\p{\uv{s} \cdot \uv{J}},\\
        ={}& -\frac{\alpha}{2}\p{\uv{s} \cdot \bar{\bm{l}}}^2
            - \alpha\p{\uv{s} \cdot \bar{\bm{l}}}
                \p{\uv{s} \cdot \bm{l}'}
            - g\p{\uv{s} \cdot \uv{J}},\\
        \approx{}& -\frac{\alpha}{2}\cos^2\theta
            - g\p{\cos \theta \cos I - \sin I \sin \theta \cos \phi}\nonumber\\
        & - \alpha \cos \theta\p{
            \cos \theta \cos i_{\rm 1f}
            - \sin i_{\rm 1f} \sin \theta \cos \p{\phi - \phi_0 - \p{g_2 -
            g_1}t}}.
\end{align}
Is this right, or is $i_{\rm 1f}$ also changing?

>>>>

It seems like we might be seeing some resonance type behaviors. For convenience,
we go to the co-rotating frame with $g_1$ where we then have
\begin{equation}
    \p{\rd{\uv{s}}{t}}_{\rm rot, 1}
        = \alpha\p{\uv{s} \cdot \uv{l}_1}\p{\uv{s} \times \uv{l}_1}
            + g_1\p{\uv{s} \cdot \uv{J}}.
\end{equation}
Here, now, the mode frequencies are $g_1 \Rightarrow 0$ and $g_2 \Rightarrow g_2
- g_1$. Decompose then $\uv{l}$ (we drop the subscript) into a mean and fluctuating
piece $\bar{\bm{l}}$ and $\bm{l}'$ (where the former is still a vector of unit
length), then:
\begin{equation}
    \p{\rd{\uv{s}}{t}}_{\rm rot, 1}
        = \alpha\p{\uv{s} \cdot \bar{\bm{l}}}\p{\uv{s} \times \bar{\bm{l}}}
            + g_1\p{\uv{s} \cdot \uv{J}}
            + \alpha\s{
                \p{\uv{s} \cdot \bar{\bm{l}}}
                \p{\uv{s} \times \bm{l}'} +
                \p{\uv{s} \cdot \bm{l}'}
                \p{\uv{s} \times \bar{\bm{l}}}
            }.
\end{equation}
This is already looking painful. Let's write down the Hamiltonian to have a
faster path to EOM in coordinates:
\begin{align}
    H_{\rm rot, 1} ={}& -\frac{\alpha}{2}\p{\uv{s} \cdot \uv{l}}^2
            - g\p{\uv{s} \cdot \uv{J}},\\
        ={}& -\frac{\alpha}{2}\p{\uv{s} \cdot \bar{\bm{l}}}^2
            - \alpha\p{\uv{s} \cdot \bar{\bm{l}}}
                \p{\uv{s} \cdot \bm{l}'}
            - g\p{\uv{s} \cdot \uv{J}},\\
        \approx{}& -\frac{\alpha}{2}\cos^2\theta
            - g\p{\cos \theta \cos I - \sin I \sin \theta \cos \phi}\nonumber\\
        & - \alpha \cos \theta\p{
            \cos \theta \cos i_{\rm 1f}
            - \sin i_{\rm 1f} \sin \theta \cos \p{\phi - \phi_0 - \p{g_2 -
            g_1}t}}.
\end{align}
Is this right, or is $i_{\rm 1f}$ also changing?

\section{06/09/21}

\subsection{Analytical Time-Dependent Hamiltonian Perturbation Theory Setup
(Abandoned)}

We came up with at least a possible way to do this perturbatively, though it
might not be useful, since we haven't included the effect of tidal dissipation
yet. I'm going to just mostly copy this from my weekly update, since I don't
want to retype the scratch work. We analyze Dong's equations, where:
\begin{align}
    \rd{\uv{s}}{t} &= \alpha\p{\uv{s} \cdot \uv{l}_1}\p{\uv{s} \times
        \uv{l}_1},\\
    \uv{l}_1 &= \Re \mathcal{I}_1\uv{x}
            + \Im \mathcal{I}_1 \uv{y}
            + \sqrt{1 - \abs{\mathcal{I}_1}^2}
            \uv{z},\label{eq:06/02/21.l1param}\\
        &= \begin{bmatrix}
            i_1 \cos (g_1t) + i_{\rm 1f} \cos\p{g_2 t + \phi_0}\\
            i_1 \sin (g_1t) + i_{\rm 1f} \sin\p{g_2 t + \phi_0}\\
            \cos\abs{\mathcal{I}_1}
        \end{bmatrix}.
\end{align}
I've taken $\phi_0 = 0$ for simplicity. We take $\alpha / g_1 = 10$
consistently. We permit $i_{\rm 1f}$ and $g_{1,2}$ to vary independently for
now, even though $i_{\rm 1f} \propto \p{g_2 - g_1}^{-1}$.

Consider the Hamiltonian
\begin{equation}
    H = - \frac{\alpha}{2}\p{\uv{s} \cdot \uv{l}}^2.
\end{equation}
Decompose $\uv{l} = \bar{\bm{l}} + \bm{l}'$ where:
\begin{align}
    \bar{\bm{l}} &= \begin{bmatrix}
        i_1 \cos (g_1t)
        i_1 \sin (g_1t)
        \cos i_1
    \end{bmatrix},\\
    \bm{l}' &\equiv \uv{l} - \bar{\bm{l}},\\
        &\approx \begin{bmatrix}
        i_{\rm 1f} \cos\p{g_2 t + \phi_0}\\
        i_{\rm 1f} \sin\p{g_2 t + \phi_0}\\
        -\frac{i_{\rm 1f}i_1}{\cos^2 i_1} \cos\p{g_2t + \phi_0}
    \end{bmatrix}.
\end{align}
Note that $l'_z \approx 0$.

We next go to the corotating frame with $g_1$ about $\uv{\jmath}$, so that
\begin{align}
    H_{\rm rot} &= - \frac{\alpha}{2}\p{\uv{s} \cdot \uv{l}}^2
        - g\p{\uv{s} \cdot \uv{\jmath}},\\
        &\equiv H_0 + H_1,\\
    H_0 &\equiv - \frac{\alpha}{2}\p{\uv{s} \cdot \bar{\bm{l}}}^2
        - g\p{\uv{s} \cdot \uv{\jmath}},\\
    H_1 &\approx -\alpha\p{\uv{s} \cdot \bar{\bm{l}}}
        \p{\uv{s} \cdot \bm{l}'} + \mathcal{O}\p{(l')^2}.
\end{align}
In the corotating frame, we find that
\begin{equation}
    \p{\bm{l}'}_{\rm rot} \approx \begin{bmatrix}
        i_{\rm 1f} \cos\p{(g_2 - g_1) t + \phi_0}\\
        i_{\rm 1f} \sin\p{(g_2 - g_1) t + \phi_0}\\
        0
    \end{bmatrix}.
\end{equation}
Thus, in the corotating frame, when $g_2 = g_1$, CS2 is simply \emph{shifted},
since $\bm{l}'_{\rm rot}$ is fixed in space. This isn't entirely realistic, as
$i_{\rm 1f} \propto \p{g_2 - g_1}^{-1}$, but also gives us a chance at using
time-dependent perturbation theory.

\textbf{NB:} Per Dong's recommendation, I'm going to mostly stay away from this,
as the dynamics are already reasonably well understood in the absence of tidal
dissipation due to resonance overlap. I suspect that there might be some other
interesting behavior, like chaotic behavior near the separatrix due to another
Melnikov's Method calculation, but it's a little bit difficult to justify doing
too much work on this when there's a lot of canonical work on the subject
already.

I do have a concern about resonance overlap: the resonances are only formally
defined in their respective corotating frames (i.e.\ they don't share resonance
angles $\phi_i = \phi_{\rm inertial} + g_it$), so I'm not exactly sure the
picture of ``don't know which resonance a point belongs to'' is entirely valid.
